---
title: "Analysis Guide: ANOVA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis Guide: ANOVA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Analysis of Variance (ANOVA)

The ANalysis Of VAriance (ANOVA) is a test of difference between two *or more* means. That is, it can functionally perform the same test as the t-test, but also can be expanded to support more means. A common example for MDRC work would be: Is the mean of some continuous variable the same between different subgroups? For example, is the average age in each region of a multi-site study the same? One thing to note is that the ANOVA returns the F-statistic, an omnibus statistic, which can be interpreted as one of the means is significantly different, but it does not specify which one. This is useful as a first test as it will identify if there is a significant different with less likelihood of multiple comparisons problems. 

## One-way ANOVA


### `stats::oneway.test()`
The `stats::oneway.test()` function is the most straightforward way to perform an ANOVA in R. The function also uses the [formula notation](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/formula), `outcome ~ predictor`, that is common in R. The `stats::oneway.test()` function also has the `var.equal=` parameter which accepts `TRUE` or `FALSE` for if variances are equal. By default, `var.equal=FALSE` is set. If `var.equal=` is set to `FALSE` then the function performs Welch's ANOVA. The below example uses the `iris` data frame which has the `Sepal.Length` (and other variables) for three species of plants. 

```{r}
oneway.test(Sepal.Length ~ Species, iris, var.equal = TRUE)
```

### `stats::aov()` with `base::summary()`
Another method for performing an ANOVA is to use the `stats::aov()` function. With the `stats::aov()` function, it is necessary to use the `base::summary()` function to evaluate the results of the test. Exploration of all values can be done with the `base::str()` function. 

```{r}
aov(Sepal.Length ~ Species, iris)

summary(aov(Sepal.Length ~ Species, iris))
```

#### `stats::aov()` with `broom` functions

The same results can be had with an easier-to-use, data frame output by using the `broom` package functions.

```{r, warning=FALSE}
library(broom)
tidy(aov(Sepal.Length ~ Species, iris))

glance(aov(Sepal.Length ~ Species, iris))

augment(aov(Sepal.Length ~ Species, iris))
```

#### `stats::aov()` Technical Notes
It might be worth noting that `stats::aov()` is a wrapper to `lm()` for fitting linear models to balanced or unbalanced experimental designs. Essentially, `stats::aov()` runs an `lm()` for each strata/group-level. 









