---
title: "Analysis Guide: T-Tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis Guide: T-Tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# T-test

The t-test is the test of equivalence between two means. A common example for MDRC work would be: Is the mean of some continuous variable the same between different treatment statuses? For example, is the average age in each level of the treatment status the same?

## `base::t.test()` function

The `base::t.test()` function is the easiest way to perform a t-test in R. The function accepts two vectors and will calculate the means and perform the t-test. In the below example, the average `Sepal.Length` for the first two species in the `iris` data set are tested for equivalence (each is 50 observations). Like with the `chisq()` function, the `t.test()` output can be viewed using the `broom::glance()` function for a data frame output. 

### Usage Notes
The t-test does not require equal sample sizes. One-sided t-tests can be performed using the `alternative=` parameter which accepts "two-sided", "less", or "greater". The function also has a `mu=` parameter for specifying a mean (for a one-sample t-test) or the null difference in means. 

```{r}
t.test(iris$Sepal.Length[1:50], 
       iris$Sepal.Length[51:100])

broom::tidy(t.test(iris$Sepal.Length[1:50], 
       iris$Sepal.Length[51:100]))

```

### Tangent

As an aside, a t-test is effectively a regression with no covariates. 
