---
title: "Analysis Guide: Chi-Square"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis Guide: Chi-Square}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Chi-Square ($\chi^2$)

The chi-square test can be used to test independence of categories in a categorical variable. The most common example at MDRC is: is some categorical outcome variable independent of treatment status? For example, is grade independent of treatment status. 

## `base::chisq()` with `broom` functions

The simplest method for performing a chi-square test is using the `base::chisq.test()` function. A basic call of the `base::chisq.test()` function will produce the expected results, but there are more results that can be had using either the `str()` function and appropriate extraction notation (`[]` or `$`) or using the `broom` package and the appropriate functions to create the desired data frames. 

```{r, warning=FALSE, message=FALSE}
chisq.test(mtcars$gear, mtcars$carb)

str(chisq.test(mtcars$gear, mtcars$carb))

chisq.test(mtcars$gear, mtcars$carb)$p.value

library(broom)
tidy(chisq.test(mtcars$gear, mtcars$carb))
glance(chisq.test(mtcars$gear, mtcars$carb))
augment(chisq.test(mtcars$gear, mtcars$carb))
```

