---
title: "mean_extract: Summary statistics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mean_extract: Summary statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Set Up

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # Disabling warnings for demonstration
  warning = FALSE
)
```

```{r setup}
library(mdrcAnalysis)
```

### Simulated data

It is the year 2050. The AI revolution is almost complete. Troves of workers across sectors & educational qualifications have been unemployed. The federal government along with the premier research agency, mdrc is partnering to evaluate job retraining programs and whether the specialized empathy & listening training (treatment) increases the chance of people being employed in new human care industries that have propped up. An initial trial has been conducted across 5 cities in multiple sites with 1-25 individuals in each training site.

The initial simulated data set is a 3 level data structure consisting of

* **City** - 5 Cities with population between 1,000 and 2,000
* **Sites** - Within each city, there are between 1-10 job training centers
* **Individuals** - Within each job training center, there are 1-25 trainees/individuals

Additionally, we are assuming that people with prior higher education has higher prior income
and have put a correlation structure around it. 
Furthermore, we think that married and single people live in distinct clusters within this data set. 
In our example below, we have set the clustering at the sites level.

```{r}
data(sim_data_robust_reg)
head(sim_data_robust_reg)
```

## Minimum Arguments

`mean_extract` requires that you pass it the data set and the treatment variable, it will return a 5 number summary along with missing values count and standard deviation for all the observations whose treatment status is not missing.

```{r}
mean_extract(
    .dataset = sim_data_robust_reg
)
```

## Subgroup

You can also specify __subgroup__ to get where metrics of different groups.

```{r}
mean_extract(
    .dataset = sim_data_robust_reg,
    .subgroup = "treatment"
)
```

## Quantiles

You can also set `quantile = TRUE` to get quantile cuts.

```{r}
mean_extract(
    .dataset = sim_data_robust_reg,
    .subgroup = "treatment",
    .quantile = TRUE
)
```

