---
title: "ttest_extract: T-Tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ttest_extract: T-Tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

library(mdrcAnalysis)
```

## Overview and Motivation

This function is effectively a wrapper around `stats::t.test` to allow staff to pass
many dependent variables, run t-tests on by groups, and easily create an extract for sharing 
or tabling.

## Sample Data

The Baseline Information Form (BIF) is a survey with a series of questions about a study participant (for example, it includes questions about their gender, date of birth, etc.). Soon after completing the BIF, study participants were randomly assigned to a program or control group. Baseline data was collected for two cohorts and provided to the data team using various file types. These synthetic BIF data are used as part of the Shoot for the Stars project in the Introduction to R and Introduction to SAS trainings. We'll be using this data in the examples below to demonstrate the ttest_extract function.

```{r, warning = FALSE, message = FALSE}
# load data set
finalBif <- bif_final
```

```{r}
head(finalBif)
```

## About the Function

This function uses `stats::t.test`. Like `stats::t.test`, this function provides 2 methods and 
will automatically select the appropriate method based on the class of the
input, although I would personally recommend sticking with passing dataframes
instead of formulas.

## Function Call Examples

### T-Test on All Numeric Variables

When passing a dataframe as the first argument, the default method is used. 
The only other required argument is the `.classvar` which is the name of the 
categorical independent variable we are testing on. The dependent variables will
by default include all numeric variables not specified by other parameters.


```{r}
ttest_extract(finalBif, .classvar = "RA_CODE")
```

This function is intended to perform and display a t-test comparison of the 
dependent variable by splitting the values of the dependent variable along 
`.classvar`. Thus `.classvar` must have exactly 2 values. If `.classvar` has any
missing values, these are filtered immediately out of the dataset. If you wish 
to perform a t-test comparison and use the existence of a variable as the 
categorical, it is recommended you create a dummy variable representing it.

### Specifying Variables

If you wish to only perform the t-test comparison on a subset of the variables
as dependents, you can explicitly specify that using  `.depvars`. `.depvar_labels` is also helpful if you have
specific labels you wish to pass along that better describe the variables.

```{r}

ttest_extract(finalBif, .classvar = "RA_CODE",
              .depvars = c("AGE", "bldiplomas_hs", "bldiplomas_ged",
                           "bldiplomas_tec", "bldiplomas_as", "bldiplomas_4yr",
                           "bldiplomas_md", "bldiplomas_non"),
              .depvar_labels = c("Age (years)", "High School Diploma",
                                 "GED", "Technical School",
                                 "Associate's Degree", "Bachelor's Degree",
                                 "Master's Degree", "No Degree"))
```

### By Group

If you wish to separate your comparisons by another variable, you can specify 
this variable with the `.byx` parameter. You can see that doing this adds a new 
column based on the `.byx` variable chosen and a UniqueID variable that combines
both the dependent variable and the `.byx` value.

```{r}

ttest_extract(finalBif, .classvar = "RA_CODE",
              .byx = "COHORT",
              .depvars = c("AGE", "bldiplomas_hs", "bldiplomas_ged",
                           "bldiplomas_tec", "bldiplomas_as", "bldiplomas_4yr",
                           "bldiplomas_md", "bldiplomas_non"),
              .depvar_labels = c("Age (years)", "High School Diploma",
                                 "GED", "Technical School",
                                 "Associate's Degree", "Bachelor's Degree",
                                 "Master's Degree", "No Degree"))  
```

## T-Test Calculcation

The other two primary arguments available to change how the t.test is calculated
are `.mu` which indicates the true difference in mean being calculated for. This
defaults to 0 which should satisfy most use cases. 

There is also `.alternative`,
which allows you to change the alternate hypothesis. This defaults to
"two.sided".

### Output to Excel

This function is largely designed around creating output in Excel. The data 
results are returned in dataframe form but will also print to an Excel if 
`.file_name` and `.file_path` are specified:

```{r}
ttest_output <- ttest_extract(
  finalBif 
  ,.classvar = "RA_CODE"
  ,.byx = "COHORT"
  ,.depvars = c("AGE", "bldiplomas_hs", "bldiplomas_ged",
                "bldiplomas_tec", "bldiplomas_as", "bldiplomas_4yr",
                "bldiplomas_md", "bldiplomas_non")
  ,.depvar_labels = c("Age (years)", "High School Diploma",
                      "GED", "Technical School",
                      "Associate's Degree", "Bachelor's Degree",
                      "Master's Degree", "No Degree")
  ,.mu = 0
  ,.alternative = "two.sided"
  # ,.output_path = "../example_extract"
  # ,.output_file = "ttest_extract_example"
  )  
```
