---
editor: 
  markdown: 
    wrap: 72
---

```{r}
devtools::load_all()
```

# Without Factors

## Without subgroup

-   No chisq
-   No missing

```{r}
crosstab_extract(.dataset = bif_final ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .chisq = F,
                 .inc_trail = F,
                 .missing = F)
```

-   Yes chisq
-   No missing

```{r}
crosstab_extract(.dataset = bif_final ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .chisq = T,
                 .inc_trail = F,
                 .missing = F)
```

-   No chisq
-   Yes missing

```{r}
crosstab_extract(.dataset = bif_final ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .chisq = F,
                 .inc_trail = F,
                 .missing = T)
```

-   Yes chisq
-   Yes missing

```{r}
crosstab_extract(.dataset = bif_final ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .chisq = T,
                 .inc_trail = F,
                 .missing = T)
```


## With subgroup 

-   No chisq
-   No missing

```{r}
crosstab_extract(.dataset = bif_final ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .subgroup = "COHORT",
                 .chisq = F,
                 .inc_trail = F,
                 .missing = F)
```

-   Yes chisq
-   No missing

```{r}
crosstab_extract(.dataset = bif_final ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .subgroup = "COHORT",
                 .chisq = T,
                 .inc_trail = F,
                 .missing = F)
```

-   No chisq
-   Yes missing

```{r}
crosstab_extract(.dataset = bif_final ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .subgroup = "COHORT",
                 .chisq = F,
                 .inc_trail = F,
                 .missing = T)
```

# With Factors

```{r}
.dataset <- bif_final |> 
  dplyr::mutate(blfemale = factor(blfemale,levels = c(0,1,2)), 
                RA_CODE  = factor(RA_CODE , levels = c("P", "C","NP"))) 
```

## Without subgroup

-   No chisq
-   No missing

```{r}
crosstab_extract(.dataset = .dataset ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .chisq = F,
                 .inc_trail = F,
                 .missing = F)
```

-   Yes chisq
-   No missing

```{r}
crosstab_extract(.dataset = .dataset ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .chisq = T,
                 .inc_trail = F,
                 .missing = F)
```

-   No chisq
-   Yes missing

```{r}
crosstab_extract(.dataset = .dataset ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .chisq = F,
                 .inc_trail = F,
                 .missing = T)
```

-   Yes chisq
-   Yes missing

```{r}
crosstab_extract(.dataset = .dataset ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .chisq = T,
                 .inc_trail = F,
                 .missing = T)
```


## With subgroup 

-   No chisq
-   No missing

```{r}
crosstab_extract(.dataset = .dataset ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .subgroup = "COHORT",
                 .chisq = F,
                 .inc_trail = F,
                 .missing = F)
```

-   Yes chisq
-   No missing

```{r}
crosstab_extract(.dataset = .dataset ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .subgroup = "COHORT",
                 .chisq = T,
                 .inc_trail = F,
                 .missing = F)
```

-   No chisq
-   Yes missing

```{r}
crosstab_extract(.dataset = .dataset ,
                 .col_var = "RA_CODE",
                 .row_vars = "blfemale",
                 .subgroup = "COHORT",
                 .chisq = F,
                 .inc_trail = F,
                 .missing = T)
```



