---
title: "Checking and Reporting on Duplicates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checking and Reporting on Duplicates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mdrcAnalysis)
library(tidyverse)
data <- read_csv(file.path(
  "/data/share/xproject/Training/Data Sources",
  "Qualtrics_TestData_210730v2.csv")) %>%
  tail(-2)
```

Checking for duplicates is a fairly simple process with tidyverse tools but 
quantifying and reporting on these duplicates is a more involved process.

`check_duplicate` performs 2 primary actions. It prints a dataframe quantifying
the amount of duplicates there are in the daata and returns the original 
dataframe with the duplicated rows flagged.

## Default Behavior

By default, `check_duplicate` only marks rows as duplicates if ALL columns are
duplicated. This is not particularly useful except in extreme cases. 

You can see here in our data that no rows are duplicated across all columns.
```{r}
check_duplicate(data)
```

## Specifying Groups

If however, we want to determine rows that are duplicated across a subset of 
columns, we can specify it as such:

```{r}
flagged_data <- check_duplicate(data, c("Q27", "Q28"))
```

Here we see a report on each combination of our grouping variables and the 
frequency that this combination occurs. This also returns an object with a flag
to allow programmers to easily filter out these duplicated cases.

```{r}
flagged_data %>% filter(DUPFLAG == 1) %>%
  select(DUPFLAG, Q27, Q28)
```


