---
title: "Creating Random Samples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating Random Samples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mdrcAnalysis)
library(tidyverse)
# Loading data
data <- readRDS("/data/share/xproject/Training/Data Sources/bif_bl.rds")
```

Creating random samples is a fairly common procedure so this function was made 
help make this process easier and with higher rproducibility.

## Default Behavior

By default, `randomsample02`creates a sample of size 50 with all columns
included. The seed used is also random but can be explicitly set with `.Seed`

```{r}
randomsample02(data)
```

## Function Arguments

The argument `.select_vars` can be used if you want your random sample to be 
column subset of your starting dataset.
```{r}
randomsample02(data,
               .select_vars = c("SAMPLEID", "blmale", "blfemale"))
```

The argument `.SampSize` is used to determine the size of the reesulting
dataset. If the value of `.SampSize` is greater than 1, then it will be used to
directly determine the resulting size. If it is less than or equal to 1, it is
treated as a proportion and is equivalent to the closest integer of `.SampSize`
multiplied by the number of rows of the original data.

```{r}
# Random sample of size 10
randomsample02(data, .SampSize = 10)
```

```{r}
# Ranndom sample of 1/10 the original size, noting that the original data had 
# 2575 rows
randomsample02(data, .SampSize = 0.1)
```

`.Replacement` defaults to `FALSE` and determines if random samples should be
taken with replacement. This also enables you to create a sample of greater size
than the original data.
```{r}
randomsample02(data, .SampSize = 3000, .Replacement = TRUE)
```

You can also group variables by passing a vector of column names to
`.groupiing_vars`. Note that `.SampSize` works on a PER GROUP basis. For the
example below, there are 3 distinct values of the `.grouping_vars` variable
"blmale". Thus, with a `.SampSize` of 10, we expect to get a sample size of 30.
Using a proportional `.SampSize` also calculates the new integer sample size 
based on the number of rows in the original data.

```{r}
randomsample02(data, .grouping_vars = "blmale", .SampSize = 10)
```
