---
title: "Datset Summary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Datset Summary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mdrcAnalysis)
library(tidyverse)
# Loading data
bif_bl <- readRDS("/data/share/xproject/Training/Data Sources/bif_bl.rds")
```

For the purposes of this demonstration, I will be using the baseline BIF data
found in "/Training/Data Sources/bif_bl.rds"

There are many ways people begin looking at a new data source before actually
processing or cleaning it. `print` or `head` are the most obvious but do not
always show a representative sample of the data and can be unwieldy with 
sufficiently many columns. 
```{r}
head(bif_bl)
```

`glimpse` is another frequently used tool and provides some improvements to a 
normal print by making the output fixed width and also dispalying the types of 
data. However, it also has the issue of not necessarily showing a the bounds of
the data.

```{r}
glimpse(bif_bl)
```

`show_ds` is designed for use in an R-Studio environment and provides a printout
summary of data intended to cover most obvious characteristics. It generates and
prints a tibble with a row for every column in the original dataset and is 
otherwise option-less because it is intended to be a simple initial summary.

```{r}
show_ds(bif_bl)
```

The summary columns are as follows:

* Variable: the name of the column
* DataType: the result of the the `typeof` call on the column vector
* DistinctVallues: The number of distinct values, including missings
* NAValues: The amount of `NA`, or missing values.
    Does not include `NaN` or `NULL`
* ValidValues: The amount of non-`NA` values.
* Mean: The mean value if numeric, `NA` if otherwise.
* Sum: The sum of all valid values if numeric, `NA` if otherwise.
* Min: The minimal numeric value, the minimal factor level, or the "1st" value
    in a typical alphanumeric sort
* Min: The maximal numeric value, the maximal factor level, or the "last" value
    in a typical alphanumeric sort
* Obs: The number of values, equal to the number of rows of the input data.
* SampleValues: A 20 character subset string concatenation of some values in the
    original data, delimited by ","
